---
title: "bigMartSalesData"
author: "dataMunglers"
date: "28 October 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* We take data from [analyticsvidhya.com](https://www.analyticsvidhya.com), Our data is [Big Mart Sales Practise Problem](https://datahack.analyticsvidhya.com/contest/practice-problem-big-mart-sales-iii/) Data, It is open comppetition now and its final date is 31 Dec 2017. We downloaded training dataset as a csv file. 

```{r eval=TRUE,echo=FALSE,include=FALSE}

setwd("C:/Users/yetkinEser/Desktop/R/datas")
bigMart<-read.csv('bigMartTrain.csv')

library(dplyr)
library(ggplot2)

```

* View the structere of bigMart with __*str*__ function


```{r}

str(bigMart)

```


* Load dplyr package and look at data types with __*glimpse*__ function 

```{r eval=TRUE,echo=FALSE,include=FALSE}

library(dplyr)

```

```{r}

glimpse(bigMart)

```

* Our dataset has 12 columns and 8523 rows
Look at all columns, Firstly factor type columns

  * *__Item_Identifier__* : Unique Product ID
  
```{r}

bigMart%>%
  summarise(n_distinct(Item_Identifier))
  
```

  * *__Item_Fat_Content__* : Whether the product is low fat or not

```{r}
  
bigMart%>%
  group_by(Item_Fat_Content)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))
  
```
  
  * *__Item_Type__* : The category to which the product belongs

```{r}
  
bigMart%>%
  group_by(Item_Type)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))

```
  
  * *__Outlet_Identifier__* : Unique Store ID

```{r}
  
bigMart%>%
  group_by(Outlet_Identifier)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))

```
  
  * *__Outlet_Size__* : The size of the store in terms of ground area covered

```{r}

bigMart%>%
  group_by(Outlet_Size)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))

```

  * *__Outlet_Location_Type__* : The type of city in which the store is located

```{r}
  
bigMart%>%
  group_by(Outlet_Location_Type)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))

```
  
  * *__Outlet_Type__* : Whether the outlet is just a grocery store or some sort of supermarket

```{r}
  
bigMart%>%
  group_by(Outlet_Type)%>%
  summarise(Count=n(),Perc=round(n()/nrow(.)*100,2))%>%
  arrange(desc(Count))

```
    
  * *__Item_Weight__* : Weight of product

```{r}
  
bigMart%>%
    summarise(is_NULL=sum(is.na(Item_Weight)==1),
              is_NOT_NULL=sum(!is.na(Item_Weight)==1)
              )
  
bigMart%>%
  filter(!is.na(Item_Weight))%>%
  summarise(
    Max=max(Item_Weight),
    Min=min(Item_Weight),
    Mean=mean(Item_Weight),
    Median=median(Item_Weight),
    QUA1=quantile(Item_Weight,1/4),
    QUA3=quantile(Item_Weight,3/4),
    IQR=IQR(Item_Weight)
  )
  
```
    
  * *__Item_Visibility__* : The % of total display area of all products in a store allocated to the particular product

```{r}
  
bigMart%>%
    summarise(is_NULL=sum(is.na(Item_Visibility)==1),
              is_NOT_NULL=sum(!is.na(Item_Visibility)==1)
              )
  
bigMart%>%
  filter(!is.na(Item_Visibility))%>%
  summarise(
    Max=max(Item_Visibility),
    Min=min(Item_Visibility),
    Mean=mean(Item_Visibility),
    Median=median(Item_Visibility),
    QUA1=quantile(Item_Visibility,1/4),
    QUA3=quantile(Item_Visibility,3/4),
    IQR=IQR(Item_Visibility)
  )
  
  
```
    
  * *__Item_MRP__* : Maximum Retail Price (list price) of the product

```{r}
  
bigMart%>%
    summarise(is_NULL=sum(is.na(Item_MRP)==1),
              is_NOT_NULL=sum(!is.na(Item_MRP)==1)
              )
  
bigMart%>%
  filter(!is.na(Item_MRP))%>%
  summarise(
    Max=max(Item_MRP),
    Min=min(Item_MRP),
    Mean=mean(Item_MRP),
    Median=median(Item_MRP),
    QUA1=quantile(Item_MRP,1/4),
    QUA3=quantile(Item_MRP,3/4),
    IQR=IQR(Item_MRP)
  )
    
```
  
  * *__Outlet_Establishment_Year__* : The year in which store was established

```{r}
  
bigMart%>%
    summarise(is_NULL=sum(is.na(Outlet_Establishment_Year)==1),
              is_NOT_NULL=sum(!is.na(Outlet_Establishment_Year)==1)
              )
  
bigMart%>%
  filter(!is.na(Outlet_Establishment_Year))%>%
  summarise(
    Max=max(Outlet_Establishment_Year),
    Min=min(Outlet_Establishment_Year),
    Mean=mean(Outlet_Establishment_Year),
    Median=median(Outlet_Establishment_Year),
    QUA1=quantile(Outlet_Establishment_Year,1/4),
    QUA3=quantile(Outlet_Establishment_Year,3/4),
    IQR=IQR(Outlet_Establishment_Year)
  )
  
```
    
  * *__Item_Outlet_Sales__* : Sales of the product in the particulat store. This is the outcome variable to be predicted.

```{r}

bigMart%>%
    summarise(is_NULL=sum(is.na(Item_Outlet_Sales)==1),
              is_NOT_NULL=sum(!is.na(Item_Outlet_Sales)==1)
              )
  
bigMart%>%
  filter(!is.na(Item_Outlet_Sales))%>%
  summarise(
    Max=max(Item_Outlet_Sales),
    Min=min(Item_Outlet_Sales),
    Mean=mean(Item_Outlet_Sales),
    Median=median(Item_Outlet_Sales),
    QUA1=quantile(Item_Outlet_Sales,1/4),
    QUA3=quantile(Item_Outlet_Sales,3/4),
    IQR=IQR(Item_Outlet_Sales)
  )
  
  
```

* Summary =)))

```{r}

summary(bigMart)

```




